Quotient Normalized Maximum Likelihood Criterion for Learning Bayesian Network Structures

### 2018-03-16 poster/qnmlPoster.tex
starting with the poster

\mode<presentation>
{
  \usetheme{hiitposter}
}
%\graphicspath{{imgs/}}
\usepackage{amsmath,amsthm, amssymb,mathrsfs}
\usepackage{color}
\usepackage{times}
\usepackage{subfigure}
\usepackage{bm}
\usepackage{url}
%\usepackage[orientation=portrait,size=a0,scale=1.4,debug]{beamerposter}        
\usepackage[orientation=portrait,width = 91 ,height = 122 ,scale=1.39,debug]{beamerposter} 
%\setbeamerfont{itemize/enumerate body}{size={\fontsize{25}{31}}}
%\setbeamerfont{itemize/enumerate subbody}{size=\footnotesize}

@@

\newcommand{\mechanism}{\mathcal{M}}
\newcommand{\dataset}{\mathcal{D}}

@@

\newcommand{\heading}[1]{\alert{\large #1}\\}
\definecolor{myPurple}{RGB}{174, 206, 195}

@@

\newcommand{\vOmega}{\bm{\Omega}}
\newcommand{\bBeta}{\bm{\beta}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\vA}{\bm{A}}
\newcommand{\vx}{\bm{x}}
\newcommand{\vy}{\bm{y}}
\definecolor{myblue}{rgb}{.85, .85, 1.0}

@@

\begin{block}{Section 1}
\begin{columns}[T]
   \begin{column}{0.3\textwidth} % first column {{{
     \heading{Subsection 1}

@@

   \begin{column}{0.005\textwidth}\linethickness{0.3ex} % separator {{{
      \color{myPurple} \line(0,1){400}
   \end{column} % }}}

   \begin{column}{0.3\textwidth} % second column {{{
     \heading{Subsection 2}
   \end{column}

   \begin{column}{0.005\textwidth}\linethickness{0.3ex} % separator {{{
      \color{myPurple} \line(0,1){400}
   \end{column} % }}}
   \begin{column}{0.3\textwidth}% third column {{{
     \heading{Subsection 3}

\begin{block}{Section 2}
\begin{columns}[T]
   \begin{column}{0.25\textwidth} % first column {{{
	\heading{Subsection 1}

@@

   \begin{column}{0.3\textwidth} % second column {{{
	\heading{Subsection 2}

@@

   \begin{column}{0.3\textwidth}% third column {{{

      \heading{Subsection 3}

@@

\begin{block}{Section 3}
  \begin{columns}[T]
    \begin{column}{0.35\textwidth}
      \heading{Subsection 1}

    \end{column}
   \begin{column}{0.005\textwidth}\linethickness{0.3ex}
      \color{myPurple} \line(0,1){800}
   \end{column} % }}}

    \begin{column}{0.23\textwidth}
    \heading{Subsection 2}

@@

    \begin{column}{0.005\textwidth}\linethickness{0.3ex} % separator {{{
      \color{myPurple} \line(0,1){800}
   \end{column} % }}}
    \begin{column}{0.3\textwidth}
    \heading{Subsection 3}

@@

\vskip-1.0ex
\begin{columns}[T]
  \begin{column}{0.63\paperwidth}
\begin{block}{References} %{{{
%\scriptsize
\tiny
{
[1] 
}
\end{block} %}}}
  \end{column}
  \begin{column}{0.33\paperwidth}
\begin{block}{Acknowledgements} %{{{
%\scriptsize
\tiny
{
..........
}
\end{block} %}}}
  \end{column}
\end{columns}
\end{frame}
\end{document}

>>>>

\mode<presentation>
{
  \usetheme{hiitposter}
}
%\graphicspath{{imgs/}}
\usepackage{amsmath,amsthm, amssymb,mathrsfs}
\usepackage{color}
\usepackage{times}
\usepackage{subfigure}
\usepackage{bm}
\usepackage{url}
\usepackage[numbers]{natbib}
\usepackage[orientation=portrait,size=a0,scale=1.4,debug]{beamerposter}        
%\usepackage[orientation=portrait,width = 91 ,height = 122 ,scale=1.39,debug]{beamerposter} 
%\setbeamerfont{itemize/enumerate body}{size={\fontsize{25}{31}}}
%\setbeamerfont{itemize/enumerate subbody}{size=\footnotesize}

@@

@@

\newcommand{\heading}[1]{\alert{\large #1}\\}
\definecolor{myPurple}{RGB}{174, 206, 195}
%\definecolor{myPurple}{RGB}{165,206,190}

@@

\definecolor{myblue}{rgb}{.85, .85, 1.0}

@@

\begin{block}{Background: Structure Learning of Bayesian Networks}
\begin{columns}[T]
   \begin{column}{0.3\textwidth} % first column {{{
     \heading{Bayesian Networks}
     \vspace*{12pt}
     \begin{itemize}
     \setlength\itemsep{1em}
     \item Provide a compact way to represent a joint distribution over a random vector $X = (X_1, \ldots , X_d)$. 
     %\item Here, each $X_i$ is a categorical random variable.
     \item Consist of:
     \begin{itemize}
     \item[1.] A Directed acyclic graph $G$ which encodes the dependencies between the components of $X$.
     \item[2.] Parameters $\theta = (\theta_1,\ldots,\theta_d)$, where $\theta_i$ denotes the parameters of the conditional distribution of $X_i$ given its parents $G_i$. 
     \end{itemize}
     \item Decomposition: $P(X \mid G, \theta) = \prod_{i=1}^dP(X_i\mid G_i,\theta_i)$
     \end{itemize}     
   \end{column}

   \begin{column}{0.005\textwidth}\linethickness{0.3ex} % separator {{{
      \color{myPurple} \line(0,1){450}
   \end{column} % }}}

   \begin{column}{0.3\textwidth} % second column {{{
     \heading{Structure Learning}
     \vspace*{12pt}
     \begin{itemize}
     \setlength\itemsep{1em}
     \item[Data:] Each $X_i$ is a categorical variable. We observe $n$ independent samples of $X$ which are collected in a data matrix $D$ of size $n\times d$. 
     \item[Goal:] We consider a \textit{score-based} approach and seek a graph $G$ that maximizes a scoring function which evaluates how well a given graph fits the observed data.
     \item Some scoring functions: BIC, BDeu and fNML. 
   	 \end{itemize}
   \end{column}

   \begin{column}{0.005\textwidth}\linethickness{0.3ex} % separator {{{
      \color{myPurple} \line(0,1){450}
   \end{column} % }}}
   \begin{column}{0.3\textwidth}% third column {{{
     \heading{Scoring Functions}
     \begin{itemize}

	 \vspace*{12pt}
	 \setlength\itemsep{1em}     
     \item[BDeu] Bayesian marginal likelihood based on Dirichlet priors. Depends on a single 
     hyperparameter $\alpha >0$ called equivalent sample size. 
     \item[BIC] Maximized likelihood with penalty $\frac{k}{2}\log n$, where $k$ is the number of free parameters in the network. 
     \item[fNML] Factorized Normalized Maximum likelihood. Maximized likelihood with penalty defined via \textit{regret} functions.
     \end{itemize}

\begin{block}{Quotient Normalized Maximum Likelihood Criterion}
\begin{columns}[T]
   \begin{column}{0.18\textwidth} % first column {{{
   \heading{Motivation}
   \vspace*{12pt}
   	\textcolor{teal}{BDeu}
	\begin{itemize}	    
    \item Is very sensitive to the choice of the hyperparameter $\alpha$ \cite{cosco.uai07}.
    \item Is not \textit{regular} \cite{Suzuki2017} (can be shown in certain situations to favour too complex models over simpler) 
    \end{itemize}
    \vspace*{12pt}
    \textcolor{teal}{fNML}
	\begin{itemize}	    
    \item Is not score equivalent: the graphs expressing the same independence statements are not scored equally.
    \item Learned structures are often rather complex, which hampers their interpretation.
    \end{itemize}

    %\vspace*{12pt} 
	%\textcolor{HIITgreen}{Figure:} Number of parameters as a function of sample size for UCI Breast Cancer  data.

@@

   \begin{column}{0.18\textwidth}
   \vphantom{\large Motivation}
   \vspace*{12pt} 
    \textcolor{teal}{BIC}
    \begin{itemize}
    \item Appears to
require large sample sizes in order to identify appropriate
structures \cite{cosco.pgm08a,Liu2012}. 
	\end{itemize}

	\begin{figure}
  	\includegraphics[width=\textwidth]{breast_cancer_npmean.pdf}
  	\caption{Number of parameters as a function of sample size for Breast Cancer (UCI) data.}
  	\end{figure}

@@

   \begin{column}{0.3\textwidth} % second column {{{
	\heading{qNML: Definition}
	\vspace*{12pt}
	\begin{itemize}
	\item We would like to choose a model $G$ maximizing	
	\begin{equation}
	P_{NML}(D;G)=\frac{P(D \mid \hat\theta , G)}{\sum_{D'}{P(D' \mid \hat  \theta,G)}},
	\end{equation}where $\hat\theta$ denotes the maximum likelihood parameters.
	\end{itemize}

@@

   \begin{column}{0.25\textwidth}% third column {{{

      \heading{qNML: Properties}

@@

\begin{block}{Experimental Results}
  \begin{columns}[T]
    \begin{column}{0.35\textwidth}
      \heading{Structure Learning}

    \end{column}
   \begin{column}{0.005\textwidth}\linethickness{0.3ex}
      \color{myPurple} \line(0,1){700}
   \end{column} % }}}

    \begin{column}{0.23\textwidth}
    \heading{Prediction}

@@

    \begin{column}{0.005\textwidth}\linethickness{0.3ex} % separator {{{
      \color{myPurple} \line(0,1){700}
   \end{column} % }}}
    \begin{column}{0.3\textwidth}
    \heading{Conclusions}

@@

\vskip-1.0ex
\begin{columns}[T]
  \begin{column}{0.63\paperwidth}
\begin{block}{References} %{{{
%\scriptsize
\tiny
{
\bibliographystyle{apalike}
\bibliography{../cosco}
}
\end{block} %}}}
  \end{column}
  \begin{column}{0.33\paperwidth}
\begin{block}{Acknowledgements} %{{{
%\scriptsize
\tiny
{
J.L., E.J. and T.R. were supported by the Academy of Finland (COIN CoE and Project \textsc{TensorML}). J.L. was supported by the DoCS doctoral programme of the University of Helsinki.
}
\end{block} %}}}
  \end{column}
\end{columns}
\end{frame}
\end{document}
